\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{times}
\usepackage{setspace}
\setlength\parskip{6pt}
\setlength\parindent{0pt}
\linespread{1.2}

% Toggle section titles on/off
\newif\ifshowsections
\showsectionsfalse % Change to \showsectionstrue to show section titles

\begin{document}

\begin{center}
    {\Large \textbf{Statement of Purpose}}
\end{center}

\ifshowsections\section*{Introduction}\fi

Emerging from a rural area in Bangladesh with limited access to technology, I cultivated a resilience
and an unwavering passion for learning. This drive propelled me to excel academically, culminating
in my pursuit of higher education at Shahjalal University of Science and Technology (SUST) in Dhaka.
Here, amidst financial and personal challenges, I unearthed a profound interest in artificial intelligence
(AI) and machine learning (ML), setting the foundation for my commitment to leveraging technology to solve
intricate real-world problems.

\ifshowsections\section*{Academic Journey}\fi

As the first in my family to pursue engineering, I approached my undergraduate studies at Shahjalal University
of Science and Technology (SUST) with ambition and determination. I consistently ranked among the top students
in my cohort, earned scholarships for academic excellence, and maintained a GPA of 3.49/4. Beyond the classroom,
I engaged in programming contests, placing 11th in the SUST National Collegiate Programming Contest (2010) with
my team, SUST 3!3, and 20th in the SUST Inter University Programming Contest (2009) with SUST SHOINIK. I also
qualified for subsequent rounds in Google Code Jam and Facebook Hacker Cup, demonstrating my competitive spirit
and problem-solving skills.

During my final year, I took \textbf{Artificial Intelligence} with \textbf{Shahidul Islam} and \textbf{Machine
Learning} with \textbf{Dr. Ruhul Amin}—both served as my course teachers at SUST and have since become assistant
professors at universities in the United States. Recognizing my perseverance and potential, they agreed to
co-supervise my undergraduate thesis. In fulfilling my commitment to their mentorship, I significantly improved
the model’s accuracy and contributed to a published paper, solidifying my dedication to advancing research
in this field.

My thesis, titled \textit{Deep-eANN: A Novel Approach for Improved Classification of Precursor microRNA Using
Artificial Neural Network Model}, ultimately reached a prediction accuracy of up to 99\%. This early research
experience highlighted the transformative potential of AI in computational biology and reinforced my passion for
developing advanced technological solutions to complex challenges.

\ifshowsections\section*{Programming Contest Experience}\fi

My commitment to excellence extended beyond academics. I actively participated in over 150 programming
contests, both online and onsite, including regional competitions. These experiences honed my analytical
skills, teamwork, and ability to perform under pressure. As a blue coder in the Codeforces, I have successfully
solved over 500 programming problems, further solidifying my expertise in algorithms and data structures.
These achievements reflect my dedication to continuous learning and my ability to tackle complex computational
challenges effectively.

\ifshowsections\section*{Personal Resilience and Sacrifices}\fi

During my second year of undergraduate studies, my father returned from overseas due to illness,
creating a need for me to support myself financially. As the second eldest among four siblings, I was
not only supporting my own living and education but also contributing to my younger siblings' expenses.
I started home tutoring students alongside my undergraduate education, earning enough to cover my
living expenses and educational costs until I graduated. This challenging period allowed me to cultivate
critical skills such as time management, leadership, and resilience. My efforts ensured that my younger
brothers could continue their education. One is now an undergraduate student, and the other is nearing the
completion of a marine degree. For over a decade, I prioritized their futures over my own aspirations,
investing all I had in their stability.

One memorable moment in my academic journey was solving a challenging dynamics problem on my own two days
after a contest where my team had struggled. This experience reinforced my belief in persistence and
individual effort, qualities I continue to rely on.

\ifshowsections\section*{Professional Experience}\fi

Spanning over a decade, my professional career has taken me across Bangladesh, Singapore, and the UAE,
allowing me to cultivate a diverse skill set in software and R\&D engineering.

In \textbf{October 2015}, I joined \textbf{AnyConnect Private Limited, Singapore}, a newly formed company by Eyeball
Networks for expanding in the field of Software as a Service (SaaS) of next-generation Internet-of-Things
(IoT) technology. At AnyConnect, I led the design and implementation of the Communication SDK,
merging my technical expertise with leadership to drive solutions that enhanced client interactions and
system efficiencies. I also supported AT\&T with on-site engineering solutions, ensuring seamless integration and
functionality of our software offerings.

Prior to that, in \textbf{October 2013}, I started working at \textbf{Eyeball Networks Inc, Bangladesh}. At
Eyeball Networks, I contributed to the development of MS-SIP Library features such as multi-video calls and
codec adaptability. These contributions were pivotal in improving communication protocols and were integral
to client success stories. My work ensured robust and scalable communication solutions, catering to diverse
client needs.

In \textbf{September 2018}, I began my tenure at Smarter AI, \textbf{Dubai, UAE}, where I currently hold the position
of Senior Software Engineer. At Smarter AI, I led the development of \textbf{Instant Streaming\textsuperscript{TM}},
a groundbreaking feature that enabled Dashcams to establish seamless connectivity within one second. This
innovation significantly enhanced the user experience and reliability of intelligent transportation systems.
Additionally, I engineered advanced traffic feature identification algorithms leveraging OpenStreetMap data,
achieving high accuracy in real-time detection of speed limits, school zones, and speed cameras. My role also
encompassed modernizing legacy systems by transitioning core SDKs to a REST API-based architecture,
improving system concurrency and performance by 40\%. Furthermore, I designed and implemented a CI/CD pipeline
that automated periodic analytics reporting and health-report analyses using Kibana and Elasticsearch,
streamlining internal processes and enhancing client reporting capabilities.

Throughout my academic and professional journey, I have been recognized for my dedication and excellence.
In addition to my programming contest accolades, I was honored as the \textbf{Heavy Lifter} of Software Releases at
Smarter AI and received multiple \textbf{Employee of the Month} awards at AnyConnect Private Limited for exceptional
performance and dedication. These recognitions underscore my ability to deliver high-quality results consistently.

\ifshowsections\section*{Research Interests}\fi

My research journey began with my undergraduate thesis and has evolved through projects exploring real-time
systems, feature detection algorithms, and optimization techniques. These efforts have deepened my technical
expertise and strengthened my commitment to advancing computer science. My skills span multiple programming
languages (C++, Python, Java) and frameworks, as well as AI/ML tools that I have applied in various professional
settings.

I am particularly interested in the intersection of AI, ML, and IoT, with a focus on developing scalable algorithms
for large-scale applications, especially in intelligent transportation and autonomous systems. Through a PhD program,
I aim to explore cutting-edge methods in deep learning and optimization, contributing meaningfully to both academia
and industry.

\ifshowsections\section*{Motivation and Future Goals}\fi

I have chosen \textbf{Sample University} for its strong reputation in interdisciplinary research and excellence
in AI and systems engineering. The work of Professor \textbf{[Faculty Name]} in \textbf{[Research Area]} closely aligns
with my research interests, and I look forward to collaborating under their mentorship. The university’s dynamic
environment and advanced resources offer the ideal platform for impactful research.

My long-term goal is to bridge academia and industry by creating innovative solutions to global challenges and
mentoring future researchers. I also remain committed to inspiring young engineers in my home community to explore
the transformative potential of technology.

\ifshowsections\section*{Conclusion}\fi

The challenges I have faced have shaped my resilience, ambition, and dedication to excellence. I believe that a
PhD from \textbf{Sample University} will equip me with the expertise to achieve my goals and make lasting
contributions to the global research community.

I look forward to joining \textbf{Sample University} as a graduate student in your esteemed department.

\begin{flushleft}
    \vspace{1.5em}
    \textbf{Nazmul Alam}
\end{flushleft}

\end{document}
